---
title: "Viz I"
output: html_document
date: "2025-09-25"
---

```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
data("weather_df")
```

Import weather data.

```{r}
ggp_weather_scatterplot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax)) +
    geom_point()

ggp_weather_scatterplot
```

Which ones are missing?

```{r}
weather_df |>
  filter(is.na(tmax))
```

# fancier scatterplots

mapping color to name variable , alpha for opacity, se removes bar.

```{r}
weather_df |>
  ggplot(aes(x = tmin,y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

where you define the aesthetics can matter. 

```{r}
weather_df |>
  ggplot(aes(x = tmin,y = tmax)) +
  geom_point(aes(color = name, size = prcp),alpha = 0.3) +
  geom_smooth(se = FALSE)
```

use faceting really quick. 

```{r}
weather_df |>
  ggplot(aes(x = tmin,y = tmax)) +
  geom_point(aes(color = name),alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df |>
  ggplot(aes(x = tmin,y = tmax)) +
  geom_point(aes(color = name),alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

More interesting scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmin,y = tmax, color = name, shape = name)) +
  geom_point(aes(color = name, size = prcp),alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

LA plot : central park data only, coverts temp to F, makes a scatterplot of min vs max temps, overlays a linear regression line 

```{r}

weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr,y = tmax_fahr)) +
  geom_point() +
  geom_smooth(se=FALSE, method = "lm")

```

## small things to know. 


